---
title: "Interactive: Tables"
author: "<br>Data Science 2"
editor: visual
format:
  revealjs:
    theme: serif
    highlight-style: tango
    slide-number: true
    embed-resources: true
    css: "../DS2slides.css"
execute:
  echo: true
  message: false
  warning: false
---

```{r setup}
#| echo: false
knitr::opts_chunk$set(attr.output = 'class="fragment"')  # doesn't work with tables
library(tidyverse)
library(kableExtra)
library(gt)
library(DT)
library(htmlwidgets)
library(reactable)
library(babynames)
```

<!-- # Announcements -->

<!-- ## Announcements (5910) -->

<!-- -   Questions from last time (GitHub in RStudio) -->

<!-- ::: fragment -->
<!-- -   Today -->

<!--     -   RTD - `patchwork/grid.arrange` (Nick) -->
<!--     -   Text Mining Project Review -->
<!--     -   Interactive - Tables -->
<!-- ::: -->

<!-- ::: fragment -->
<!-- -   Upcoming -->

<!--     -   Thursday - RTD `scales/fonts` (Valerija), Interactive Tables/Visualizations -->
<!--     -   Tue 3/24 -->
<!--         -   RTDs - Python Integration (Jake), Quarto Presentations (Chris) -->
<!--         -   Manage on your own -->
<!--         -   Recorded -->
<!--     -   No Class - Thu 3/26, Tue 3/30 (Extra Tue), Thu 4/2 (Easter Break). Work on... -->
<!--         -   Homework #8 -->
<!--         -   Data Acquisition / Wrangling and Visualizations (non-Interactive) for Presentations -->
<!-- ::: -->

<!-- ## Registration - Fall 2026 -->

<!-- -   Statistics -->

<!--     -   Stat 4444: Applied Time Series (Alcantara) -->
<!--     -   Stat 4480/8480: Data Mining (Zhang) -->
<!--     -   Stat 5700: Probability (Guthrie) -->
<!--     -   Stat 8450: Longitudinal Data Analysis (Bernhardt) -->
<!--     -   Stat 8462: Stochastic Modeling (Frey) -->

<!-- -   Computer Science -->

<!--     -   CSC 3010: Overview of Cybersecurity (Anderson) -->
<!--     -   CSC 4480: Principles of Database Systems (Ebelechukwu/Scales) -->
<!--     -   CSC 4505: Applied Machine Learning (DeBenedetto) -->

<!-- ## DataFest Philly 2026 -->

<!-- <br><br> -->

<!-- -   [Register](https://docs.google.com/forms/d/e/1FAIpQLSfD2lFnLrgfQhgiws_aOXe8EW4LJYN2raERRQInzAvFwEwxMw/viewform) as teams of 2-5 students or as individual by March 28 to guarantee spot -->
<!-- -   Details at <https://bit.ly/datafestphilly> -->
<!-- -   Last year we had 15 teams from 10 local colleges/universities -->
<!-- -   20 consultants / judges from various industries -->
<!-- -   Free food and prizes throughout the weekend -->

<!-- # Announcements -->

<!-- ## Announcements (8796) -->

<!-- -   Questions from last time - SQL / GitHub -->

<!-- ::: fragment -->
<!-- -   Today -->

<!--     -   RTDs - `gganimate` (Rob), `ggpointdensity`/`ggridges` (Selin), `esquisse` (CC) -->
<!--     -   Text Mining Project Review -->
<!--     -   GitHub in RStudio -->
<!--     -   Interactive - Tables -->
<!-- ::: -->

<!-- ::: fragment -->
<!-- -   Upcoming -->

<!--     -   Tue 3/24 -->
<!--         -   RTDs - Madison, Joe, Will M, (break), Jerry, Donovan, Zixu -->
<!--         -   Manage on your own -->
<!--         -   Recorded -->
<!--     -   No Class - Tue 3/30 (Extra Tue) -->
<!--     -   Work on... -->
<!--         -   Homework #8 -->
<!--         -   Data Acquisition / Wrangling and Visualizations (non-Interactive) for Presentations -->
<!-- ::: -->


# Interactive Data Visualizations

1.  Data Tables (`DT`)
2.  Visualizations (`plotly`)
3.  Dashboards (`Shiny`)

## Interactive Data Viz - Overview

::: incremental
-   JavaScript applications are popular

-   D3.js (“data-driven documents”) is a Javascript library

    -   Allows for interactive graphcs using Scalable Vector Graphics (SVG), HTML5, and Cascading Style Sheets (CSS)
    -   Pushes the burden of the work to display the data onto the client side
    -   ...ALL of the data is stored in the html file!
    -   integrates nicely into R using `htmlwidgets` package
:::

# Shooting Data (Recall)

```{r}
shoot_file <- "https://raw.githubusercontent.com/washingtonpost/data-police-shootings/master/v2/fatal-police-shootings-data.csv"
shootings <- read_csv(shoot_file)
```

The Washington Post compiled a database of every fatal shooting in the United States by a police officer in the line of duty since January 1, 2015.
Of note, there have been 2,664 shootings as of the last update (which was on Sept 19, 2017).
The Post is tracking more than a dozen details about each killing — including the race of the deceased, the circumstances of the shooting, and whether the person was armed — by culling local news reports and monitoring independent databases, such as Killed by Police and Fatal Encounters.
In some cases, The Post conducted additional reporting.
The Post is documenting only shootings in which a police officer, while on duty, shot and killed a civilian — circumstances that most closely parallel the 2014 killing of Michael Brown in Ferguson, Mo.
The Post is not tracking deaths of people in custody, fatal shootings by off-duty officers, or deaths in which police gunfire did not kill the individual.
The FBI and the Centers for Disease Control and Prevention log fatal shootings by police, but officials acknowledge that their data is incomplete.

# Tables

## Table - Default

```{r}
#| code-line-numbers: "2|3|4-5|1|6|"
shoot_sum <- 
  shootings |> 
  group_by(state) |> 
  summarize(numshoot = n(),
            pct_gun = round(100*mean(armed_with == "gun",na.rm=T),1))
shoot_sum
```

## Table - gt

`gt::gt`

```{r}
shoot_sum |> gt()
```

## Table - Kable

`kableExtra::kbl` and `kableExtra::kable_classic_2`

```{r}
shoot_sum |> kbl() |> kable_classic_2()
```

## revealjs - Output / Scrollable Tables

<br><br>

Note that `attr.output = 'class="fragment"'` doesn't do anything here.

<br>

::: fragment
Add to slide title:<br> `## Table - Kable {.scroll-table}`

Add to css:<br> `.scroll-table {`<br> `overflow-x: auto;`<br> `overflow-y: auto;`<br> `}`
:::

## Table - Kable with `{.scroll-table}` {.scroll-table}

```{r}
shoot_sum |> kbl() |> kable_classic_2()
```

# Interactive Tables with `DT::datatable`

## Overview

::: incremental
-   Interfaces with JavaScript's DataTables
-   Makes interactive tables searchable, sortable, pageable
-   Particularly useful when deploying online
:::

## Base datatable

`DT::datatable`

```{r}
shoot_sum |> datatable()
```

## revealjs - Smaller datatables

<br>

Add to slide title: `## Table - Kable {.small-dt}`

Add to css:<br> `.small-dt .dataTables_wrapper` {<br> `font-size: 0.7em;`<br> `}`

::: fragment
<br>

For reactable below:

Add to css:<br> `.small-reactable .reactable` {<br> `font-size: 0.7em;`<br> `}`
:::

## Base datatable with `{.small-dt}` {.small-dt}

```{r}
shoot_sum |> datatable()
```

::: fragment
See Quarto File and Explore...
:::

## Better datatable {.small-dt}

```{r}
shoot_sum |> 
  datatable(
    rownames = F,  # removes numbers on left
    colnames = c("State", "# Shootings", "% of time using gun"),
    caption = "Shootings per State")
```

## Scrollable datatable {.small-dt}

```{r}
shoot_sum |> 
  datatable(
    options = list(
      scrollX = TRUE,   # horizontal scrolling
      scrollY = "300px", # vertical scrolling
      pageLength = 10  # limits page length
    ))
```

## Filterable datatable {.small-dt}

```{r}
shoot_sum |> datatable(filter = "top")
```


## `reactable` package {.small-reactable}

Alternative to `DT`.  Often better for Shiny Apps or large tables.

```{r}
shoot_sum |> reactable()
```

## `reactable` package = `datatable` {.small-reactable}

```{r}
shoot_sum |> reactable(
  searchable = TRUE,
  filterable = TRUE,
  pagination = TRUE,
  defaultPageSize = 10
)
```

# Saving HTML objects

Useful for many different types (including `plotly`)

```{r}
#| eval: false
#| code-line-numbers: "10"
shoot_sum |> 
  datatable(
    rownames = F,  # removes numbers on left
    colnames = c("State", "# Shootings", "% of time using gun"),
    caption = "Shootings per State",
    options = list(
      scrollX = TRUE,   # horizontal scrolling
      scrollY = "300px" # vertical scrolling
    )) |> 
  htmlwidgets::saveWidget("shooting table.html")
```

# Your Turn - DT

Using the `babynames` dataset (from the [babynames](https://cran.r-project.org/web/packages/babynames/index.html) package) to create a well-labeled, interactive data table that shows the following for each *name-sex* dyad:

-   Number of babies born of the *name-sex*
-   Percent of years in the data in which that *name-sex* appeared, rounded to one decimal place
-   Across all years, the rank of that *name* within that *sex* (i.e. Mary and James both get a rank of 1 as the most popular female and male names, respectively)

Then, **use this table** to...
(note, this is not reproducible without a log):

1.  Show the rank(s) for your name (for both males and females, if they both exist) (if your name is not on the list, choose another name to explore)

2.  Among all the names with the lowest proportion of years included, which name has the highest rank?

```{r}
#| eval: false
#| echo: false
num_year <- babynames |> distinct(year) |> nrow()

babynames |> 
  group_by(name, sex) |> 
  summarize(
    num = sum(n),
    prop = n() / num_year,
    .groups = "drop"
    ) |> 
  group_by(sex) |> 
  mutate(rank = rank(-num) |> scales::comma(),  # same as rank(desc(num))
         num = scales::comma(num),
         prop = scales::percent(prop,0.1)) |>   
  ungroup() |> 
  datatable(rownames = F,
            colnames = c("Name", "Sex", "Total Births", "Proportion of Years", 
                         "Rank (within Sex)"),
            caption = "Babynames Data",
            filter = "top")

# Michael (male 4, female 926)
# Christop (male 4316.5)
```


```{r}
#| eval: false
#| echo: false
num_year <- babynames |> distinct(year) |> nrow()

babynames |> 
  group_by(name, sex) |> 
  summarize(
    num = sum(n),
    prop = n() / num_year,
    .groups = "drop"
    ) |> 
  group_by(sex) |> 
  mutate(rank = rank(-num) |> scales::comma(),  # same as rank(desc(num))
         num = scales::comma(num),
         prop = scales::percent(prop,0.1)) |>   
  ungroup() |> 
  reactable(filterable = TRUE,
            searchable = TRUE)
```

# Resources

-   DT - [CRAN](https://cran.r-project.org/web/packages/DT/) and [RStudio guide](https://rstudio.github.io/DT/)
-   Other packages on CRAN - [reactable](https://cran.r-project.org/web/packages/reactable), [htmlwidgets](https://cran.r-project.org/web/packages/htmlwidgets), [babynames](https://cran.r-project.org/web/packages/babynames/index.html)
-   About [Javascript](https://en.wikipedia.org/wiki/JavaScript) and [D3.js](https://en.wikipedia.org/wiki/D3.js) (Wikipedia)
-   [Table printing demo](https://debruine.github.io/quarto_demo/table.html)


<!-- Start with Your Turn UG on 3/19 (they started on it) -->