---
title: "Homework 5"
author: "Josh Mazurkiewicz"
date: February 17, 2026
date-format: long
editor: source
format:
  html:
    embed-resources: true
execute:
  echo: true
  message: false
  warning: false
---

```{r setup, echo = FALSE}
library(tidyverse)
library(schrute)
library(tidytext)
```


```{r}
office_data <- theoffice %>% 
    mutate(lines = str_to_lower(text)) %>% 
    select(-director, -episode_name, -writer, -imdb_rating, - total_votes, -air_date)
```

# 1)

```{r}
office_data %>%
    pull(lines) %>%
    str_count("that's what she said") %>% 
    sum()

office_data %>%
    pull(lines) %>%
    str_detect("that's what she said") %>% 
    sum()
```
"That's what she said" is explicitly said 35 times. However, there are only 33 lines that contain that explicit expression. This is because in line 20370, Pam says the phrase 3 times.




# 2)

```{r}
office_4gram <- office_data %>% 
    unnest_tokens(words_4, lines, token = "ngrams", n = 4)

office_4gram %>% 
    head(10)
```



```{r}
office_4gram %>% 
    filter(words_4 == "that's what she said") %>% 
    count()
```


This matches the first part for number of times the phrase itself is said.


# 3) 


```{r}
office_4gram %>% 
    filter(str_detect(words_4, "^that's what she")) %>% 
    filter(words_4 != "that's what she said")
```


There are three additional phrases that are selected here, but only two that I would include in the context of "That's what she said:" 1902, 18385


# 4) 

```{r}
office_4gram %>%
    filter(str_detect(words_4, "what she said"), 
           !str_detect(words_4, "that's what she said")) %>%
    mutate(before = str_detect(words_4, "that's what she said")) %>%
    select(-season, -episode, -text_w_direction)
```

It does not appear, after inspecting the instances, that any additional phrases here should be included in context.



# 5)


```{r}
office_data %>%
    filter(
        str_detect(lines, "that"),
        str_detect(lines, "what"),
        str_detect(lines, "she"),
        str_detect(lines, "said"),
        !str_detect(lines, "that's what she said")) %>% 
    pull(lines)
```

There do not appear to be any other lines that I would consider, given they have "that", "what", "she", and "said" in them.


# 6)

```{r}
that_what_X <- office_4gram %>%
    filter(str_detect(words_4, "^that's what .* said$")) %>%
    separate(words_4, into = c("that", "what", "X", "said"), sep = " ") %>% 
    count(X, sort = TRUE)
```


```{r}
ggplot(data = that_what_X) + 
    geom_col(aes(x = reorder(X, -n), y = n, fill = n), show.legend = FALSE) + 
    theme_minimal() +
    ggtitle("Distribution of \"that's what X said\"") + 
    xlab("X") + 
    ylab("Count")
```






Investigating the context of what each "that's what X said" was to see if I would include it.
```{r}
x_said_lines0 <- office_4gram %>%
    filter(str_detect(words_4, "^that's what .* said$")) %>%
    separate(words_4, into = c("that", "what", "X", "said"), sep = " ") %>% 
    select(-that, -what, -said) %>% 
    filter(X != "she") %>% 
    pull(index)

x_said_lines <- c(x_said_lines0, x_said_lines0 - 1, x_said_lines0 + 1)


office_data %>% 
    filter(index %in% x_said_lines)
```


Include Lines in context: 16543, 22520, 36518



# 7)

```{r}
office_data <- office_data %>% 
    mutate(twss_type = as.factor(ifelse(
        str_count(lines, "that's what she said"), 1, ifelse(
            index %in% c(1902, 18385), 3, ifelse(
                index %in% c(), 4, ifelse(
                    index %in% c(), 5, ifelse(
                        index %in% c(16543, 22520, 36518), 6, 0)))))))
```






# 8)

```{r}
office_data %>% 
    count(twss_type) %>%
    ggplot() + 
    geom_col(aes(x = reorder(twss_type, -n), y = n, fill = n), show.legend = FALSE) + 
    theme_minimal() +
    ggtitle("Distribution of Identification of \"that's what she said\"") + 
    xlab("Question Identified (0 if line does not include twss)") + 
    ylab("Count of Lines")
```



```{r}
office_data %>% 
    count(twss_type) %>%
    ggplot() + 
    aes(x = twss_type, y = n) +
    geom_col(fill = "red", color = "black", show.legend = FALSE) + 
    theme_minimal() +
    ggtitle("Distribution of Identification of \"that's what she said\"") + 
    xlab("Question Identified (0 if line did not include twss)") + 
    ylab("Count of Lines") + 
    coord_cartesian(ylim = c(0, 50)) + 
    geom_text(aes(y = 2, label = paste0(n, " lines")), vjust = -1)
```






